TODO: A Survey of StarCraft AI Techniques

### 2012-03-22 survey.tex
Unifying section 'Existing work'

{\color{blue}
Introduce the game plus screenshot. Not too lengthy.
}
{\color{ForestGreen}
Here it is a brief description, I don't think we need a screenshot.
}
Starcraft: Brood War is a popular RTS game released in 1998 by Blizzard Entertainment. Starcraft is set in a science-fiction based world where the player must choose one of the three races: Terran, Protoss or Zerg. The good work done by the people of Blizzard makes this game one of the most extremely well-balanced RTS games ever created.
\begin{itemize}
	\item Terrans, human exiled from Earth, provide units that are versatile and flexible giving a balanced option between Protoss and Zergs.
	\item Protoss units have lengthy and expensive manufacturing process, but they are strong and resistant. These conditions makes players follow a strategy of quality over quantity.
	\item Zergs, the insectoid race, units are cheap and weak. They can be produced fast, encouraging players to overwhelm their opponents with sheer numbers.
\end{itemize}

@@

{\color{blue}
Include here the literature review of existing work on RTS Game AI. We could also compare with what is being used in industry, and use the real AI of Starcraft as an example (Bob Fitch's talk in AIIDE provides enough material for that). I have included here as subsections the two blurbs of text we had for integration of domain knowledge and reasoning with uncertainty, since I guess they correspond here.
}

\subsection{Micro-management}

Some interesting uses of reinforcement learning (RL) \cite{Sutton} to RTS research are in the domain of \textit{micro-management} (group and individual units control): \cite{Marthi05} employs concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. \cite{Madeira06} advocates the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. In real game setups, the actions spaces to explore is gigantic, it requires to use the structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. There are other was to learn parameters of an underlying model, for instance \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. The difficulty to work with multi-scale goals and plans is handled directly by case-based reasoning (CBR), which has been adapted for micro-management with continuous action models \cite{Molineaux08}, an integrated RL/CBR algorithm using continous models, or with hybrid CBR/RL transfer learning \cite{CBR-RL}. Reactive planning (for behavior languages) \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration.

Potential fields \cite{Hagelback08,Hagelback09} are a powerful way to specify complex (with the combination of several fields, requiring a parameter-tuning method) behavior policies and are used in modern commercial games and RTS AI are using derivative of it (Supreme Commander flowfield). Self-organizing-maps (SOM) learned with evolutionary algorithms provide some way to learn (realistic) parameters for RTS units behavior \cite{teamCompositionRTS}. Bayesian programming has been applied to inverse fusion of the sensory inputs of the units \cite{SynnaeveMicroCig11}, which subsumes potential fields, allowing for integration of tactical goals directly in micro-management. There are also some cognitive approaches to RTS AI \cite{SORTS}, in which there is a human-like attention model for micro-management.

\subsection{Opponent modeling, strategy prediction}

There are related works in the domains of opponent modeling \cite{HsiehS08,SchaddBS07,OBRecog}. The main methods used to these ends are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}. There are precedent works of Bayesian plan recognition \cite{BMPR}, even in games with Albrecht \textit{et al.} \cite{BayesianRecog} using dynamic Bayesian networks to recognize a user's plan in a multi-player dungeon adventure.

\cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR and expert demonstrations on Wargus. They update their previous CPB approach by using ``situation assessment'' for plan retrieval. They improve the speed of CPB by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences. Strategies are choices of building construction order in their model.

\cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behaviour and they applied their work to the Spring RTS (Total Annihilation clone). \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN. %\cite{Madeira06} 
\cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning on StarCraft replays. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). \cite{SynnaeveOpeningCig11} continued on \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on clustering) model to learn from replays and predict openings (early game strategies) from StarCraft replays, while \cite{SynnaeveAIIDE11} presents a whole game Bayesian tech-tree predictor based on unserpervised learning from replays.

@@

\subsection{Terrain Analysis}
{\color{ForestGreen}
Terrain analysis supplies the AI with chunks of abstract information about the map in order to help in making decisions. Terrain analysis can provide automatic detection of regions, choke points and other areas of interest. This analysis is usually performed off-line, in order to save processing time during gameplay.

The final goal of terrain analysis is having all the useful tactical information of the map. To do this we can use different algorithms, one of the most used, coming from computer vision, being \textit{skeletons by influence zones}, also known as \textit{Voronoi diagrams}\cite{Perkins_2010}. Through this kind of diagrams we will be able to generate a graph with the centroid of the regions and the choke points of each region.

Another common technique is the utilization of an influence map to detect and mark relevant zones of the map as resources location.
}

\subsection{Pathfinding}
{\color{ForestGreen}
The most common pathfinding algorithm is A*, but the big problem of this algorithm is the time and  memory consumption, a show-stopper in a real-time environment.

** There are a lot of new pathfinding algorithms. Do we need to review all of them? **
}

\subsection{Group behavior}
{\color{ForestGreen}
Craig Reynols introduced algorithmic steering behaviors to simulate the aggregate motion of swarms using procedural animation\cite{Reynolds_1999}. His model describes how a system of multiple agents , each steering according to simple local rules, can produce the collective movement found in flocks of birds. With this behaviours we are able to simulate intelligent squad movements and do tactical moves to flank the opponent in a RTS game\cite{Danielsiek_2008}, improving the survival of our units.

A good alternative to Flocking is using Reciprocal Velocity Obstacles (RVO)\cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions.With this algorithm we can move a large squad crossing another squad and avoid all possible collisions between units. This type of techniques are already applied in commercial RTS games like Starcraft 2 or Supreme Commander 2 with great success.

Another popular technique is Potential Fields. basically, Potential Fields work like a charged particle moving through a magnetic field. The process of tuning and optimizing the parameters of potential field charges can be time reduced applying learning algorithms such as Q-Learning\cite{Liu_2008}. Another common issue we have to deal with when using potential fields is the local optima problem that can stuck the unit away from the goal.
Some examples of the utilities of potential files in RTS games are avoiding obstacles (navigation), avoiding opponent fire, or staying at maximum shooting distance\cite{Hagelback08}\cite{Hagelback09}.
}

\subsection{Decision Making}
{\color{ForestGreen}
The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

A good alternative are Finite State Machine. The idea behind FSM is to decompose an object's behavior into easily manageable "chunks" or states, such as "attacking an enemy", "gathering resources" or "repairing" and establish the conditions that trigger transitions between them.
}

\subsection{Opponent modelling}
{\color{ForestGreen}
Opponent modelling is the process of keeping track of what the enemy is doing during a game in order to estimate the probability of it using a specific strategy. The key point here is to predict what units the enemy will build, when he will build them, when and where he will attack and so on. This will give us information about what kind of opponent it is allowing us to adapt our strategy and select the best counter.

Science we are dealing with imperfect information, the best solution is to build a Bayesian model which can deal perfectly with the uncertainty. The difficult part is building that model. Some approaches use traces of game logs to analyze the state of the world, but those traces have all the information of the game, so we have to add some noise to emulate the real use of the model in a non-deterministic world.

A Threat Map or Influence Map is a way to store relevant information about the opponent and keep track of the conflict areas or the secure ones. Statical threat maps can be designed where the threat is always "real" or concepts like heat map over time can be applied to make predictions.
}

@@

>>>>

Starcraft: Brood War is a popular RTS game released in 1998 by Blizzard Entertainment. Starcraft is set in a science-fiction based world where the player must choose one of the three races: Terran, Protoss or Zerg. The good work done by the people of Blizzard makes this game one of the most extremely well-balanced RTS games ever created.
\begin{itemize}
	\item Terrans, human exiled from Earth, provide units that are versatile and flexible giving a balanced option between Protoss and Zergs.
	\item Protoss units have lengthy and expensive manufacturing process, but they are strong and resistant. These conditions makes players follow a strategy of quality over quantity.
	\item Zergs, the insectoid race, units are cheap and weak. They can be produced fast, encouraging players to overwhelm their opponents with sheer numbers.
\end{itemize}

@@

\subsection{Perception}
Every AI needs to observe the world in order to analyze it and take decisions. In RTS games this task isn't trivial since we have to deal with imperfect information.

\textbf{Spatial}

Terrain analysis supplies the AI with chunks of abstract information about the map in order to help in making decisions. This analysis is usually performed off-line, in order to save processing time during gameplay.

Pottinger \cite{Pottinger00} described \emph{BANG!} engine implemented by Ensemble Studios for the game Age of Empires 2. This engine provides terrain analysis functionalities to the game using influence maps and areas with connectivity information. Later works \cite{Forbus01} showed the importance to have qualitative spatial information, and some others \cite{Karavelas04}\cite{Hale08} focused on spatial decomposition. The most recent work \cite{Perkins10} presented an algorithm using Voronoi diagrams to detect regions and relevant choke points.

\textbf{Temporal}

In \cite{WeberAIIDE11} a particle model is used to tracks opponent units under fog of war over time.

{\color{ForestGreen}
Anyone did anything here? I think only Skynet uses temporal reasoning...
Maybe heat maps?? but I can't find any reference...
}

\textbf{Enemy}

Opponent modeling is the process of keeping track of what the enemy is doing during a game in order to estimate the probability of it using a specific strategy.

The main methods used to these ends are case-based reasoning (CBR) and planning or plan recognition \cite{LTW,CBR_Planning,OntanonCBR,HTNPlanning,Ramirez}.\cite{LTW} used CBR to perform dynamic plan retrieval extracted from domain knowledge in Wargus (Warcraft II clone). \cite{CBR_Planning} base their real-time case-based planning (CBP) system on a plan dependency graph which is learned from human demonstration. In \cite{OntanonCBR,PlanRetrieval}, they use CBR and expert demonstrations on Wargus. They update their previous CPB approach by using ``situation assessment'' for plan retrieval. They improve the speed of CPB by using a decision tree to select relevant features. Hsieh and Sun \cite{HsiehS08} based their work on \cite{LTW} CBR model and used StarCraft replays to construct states and building sequences. Strategies are choices of building construction order in their model.

\cite{SchaddBS07} describe opponent modeling through hierarchically structured models of the opponent behavior and they applied their work to the Spring RTS (Total Annihilation clone). \cite{HTNPlanning} use hierarchical task networks (HTN) to model strategies in a first person shooter with the goal to use HTN planners. \cite{OBRecog} improve the probabilistic hostile agent task tracker (PHATT \cite{PHATT}, a simulated HMM for plan recognition) by encoding strategies as HTN. %\cite{Madeira06} 
\cite{WeberCig09} proposed ``a data mining approach to strategy prediction'' and performed supervised learning on StarCraft replays. \cite{HMMstrat_RTS_AIIDE11} used HMM to learn the transition probabilities of sequences of constructions and kept the most probables to produce probabilistic behavior models (in StarCraft). \cite{SynnaeveOpeningCig11} continued on \cite{WeberCig09} and used a Bayesian semi-supervised (labeling based on clustering) model to learn from replays and predict openings (early game strategies) from StarCraft replays, while \cite{SynnaeveAIIDE11} presents a whole game Bayesian tech-tree predictor based on unsupervised learning from replays.

Most commercial games avoid gather this information making their AI omniscient and using perfect information all the time. But sometimes they simulate some scouting tasks as Bob Fitch described in his AIIDE 2011 keynote for the game series WarCraft and StarCraft I.

{\color{ForestGreen}
Nobody had described an efficient way to scout the opponent... better opponent information better opponent modeling...
}

\textbf{Ally}

{\color{ForestGreen}
Anyone did anything here?
}

\subsection{Strategy}

The most common decision making are Decision Trees. Some of the advantages of the decision trees are: simple to understand and interpret acting like a white box where we can explain why a decision is taken, and they are compatible with other decision techniques. For complex decisions we have many algorithms to generate an optimum decision tree like ID3 or C4.5, depends on the problem we will use one or another.

A good alternative are Finite State Machine. The idea behind FSM is to decompose an object's behavior into easily manageable "chunks" or states, such as "attacking an enemy", "gathering resources" or "repairing" and establish the conditions that trigger transitions between them.

\subsection{Giving orders}

{\color{ForestGreen}
Explain centralized models (like blackboard) and distributed/hierarchically models.
}

\subsection{Executing orders}

\textbf{Pathfinding}

The most common pathfinding algorithm is A*, but the big problem of this algorithm is the time and  memory consumption, a show-stopper in a real-time environment.

{\color{ForestGreen}
There are a lot of new pathfinding algorithms. Do we need to review all of them?
}

\textbf{Group behavior}

Craig Reynols introduced algorithmic steering behaviors to simulate the aggregate motion of swarms using procedural animation\cite{Reynolds_1999}. His model describes how a system of multiple agents , each steering according to simple local rules, can produce the collective movement found in flocks of birds. With this behaviours we are able to simulate intelligent squad movements and do tactical moves to flank the opponent in a RTS game\cite{Danielsiek_2008}, improving the survival of our units.

A good alternative to Flocking is using Reciprocal Velocity Obstacles (RVO)\cite{Berg_2008}, since it is easier to control the behavior and they guarantee to prevent collisions. With this algorithm we can move a large squad crossing another squad and avoid all possible collisions between units. This type of techniques are already applied in commercial RTS games like Starcraft 2 or Supreme Commander 2 with great success.

Another popular technique is Potential Fields. basically, Potential Fields work like a charged particle moving through a magnetic field. The process of tuning and optimizing the parameters of potential field charges can be time reduced applying learning algorithms such as Q-Learning\cite{Liu_2008}. Another common issue we have to deal with when using potential fields is the local optima problem that can stuck the unit away from the goal. Some examples of the utilities of potential files in RTS games are avoiding obstacles (navigation), avoiding opponent fire, or staying at maximum shooting distance\cite{Hagelback08,Hagelback09}.

There are some interesting uses of reinforcement learning (RL) \cite{Sutton}: \cite{Marthi05} employs concurrent hierarchical Q-learning (units Q-functions are combined at the group level) RL to efficiently control units in a ``one robot with multiple effectors'' fashion. \cite{Madeira06} advocates the use of prior domain knowledge to allow faster RL learning and applied their work on a turn-based strategy game. In real game setups, the actions spaces to explore is gigantic, it requires to use the structure of the game in a partial program (or a partial Markov decision process) and a shape function (or a heuristic) \cite{Marthi05}. There are other was to learn parameters of an underlying model, for instance \cite{GA} used evolutionary learning techniques, but face the same problem of dimensionality. The difficulty to work with multi-scale goals and plans is handled directly by case-based reasoning (CBR), which has been adapted for units behavior with continuous action models \cite{Molineaux08}, an integrated RL/CBR algorithm using continuous models, or with hybrid CBR/RL transfer learning \cite{CBR-RL}. Reactive planning (for behavior languages) \cite{WeberCig10}, a decompositional planning similar to hierarchical task networks \cite{HTNPlanning}, allows for plans to be changed at different granularity levels and so for multi-scale (hierarchical) goals integration.

Self-organizing-maps (SOM) learned with evolutionary algorithms provide some way to learn (realistic) parameters for RTS units behavior \cite{teamCompositionRTS}. Bayesian programming has been applied to inverse fusion of the sensory inputs of the units \cite{SynnaeveMicroCig11}, which subsumes potential fields, allowing for integration of tactical goals directly in micro-management. There are also some cognitive approaches to RTS AI \cite{SORTS}, in which there is a human-like attention model for micro-management.

@@

@@

In his AIIDE 2010 keynote, Chris Jurnet described the technique XXX, implemented by [COMPANY] in the game YYY \cite{gameurl}
